---
title: "TidyTemplate for Tornados data"
date: 2023-07-10
output: html_document
---

# Overview

This is the first part of the RStudio Lab series. In this series we will be talking about how to use RStudio for Data Analytics.

Let's set some chunks options and load necessary libraries. In this first week series we use the #TidyTuesady data downloaded from the {tidytuesdayR} package.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```


# How to load the data

To download the weekly data and make available in the `tt` object:
```{r Load}
tt <- tt_load("2023-05-16")
```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}
tt 
```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse) %>% dim
```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}
tornados <- tt$tornados
tornados%>%head
```


```{r}
tornados$mag %>% summary()
```


```{r Wrangle2}
my_sts <- tornados %>%
  select(yr,mo,st,mag) %>% # dim
  drop_na() %>% # dim
 count(st,sort=T) %>%
  slice(1:5) %>%
  distinct(st)%>%
  unlist()
```



```{r}
my_tornados_sts <- tornados %>%
  select(yr,mo,st,mag) %>% # dim
  drop_na() %>%
  filter(st %in% my_sts) %>%
  group_by(st,yr)%>%
  reframe(av_mag=mean(mag))

my_tornados_sts %>%head
```


# Visualize

Using our processed dataset, we can create our unique visualization.

```{r eval=FALSE}
?geom_smooth
```

```{r Visualize}

my_tornados_sts %>%
  ggplot(aes(x=yr,y=av_mag,group=st,fill=st),color="grey")+
  geom_point(shape=21,stroke=0.5)+
  geom_smooth(aes(color=st),se=F)+
  labs(title="Tornados in five US countries from 1950",
       x="Year(Time)", y= "Magnitude",
       caption="DataSource: NOAA's National Weather Service Storm Prediction Center")+
  ggthemes::theme_fivethirtyeight()
  
  
```

# Save Image

Save the image of our plot for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "images/01_scatterplot_Tornados.png",
  device = "png")

```
