---
title: "TidyTuesday - Tornados map"
date: 2023-07-11
output: html_document
---

# Overview

Tornadoes are a type of violent rotating column of air that extends from a thunderstorm to the ground. They are capable of causing widespread damage and can occur anywhere in the world, but are most common in the United States. Tornadoes are typically characterized by their funnel shape and high winds, which can exceed 300 miles per hour (480 kilometers per hour). They often appear as a dark, rotating cloud, with a whistling or roaring sound. Tornadoes can vary in size and strength, with some lasting only a few seconds and causing minimal damage, while others can last for hours and leave a trail of destruction in their wake.

The dataset provided was derived from a dataset produced by NOAA's Storm Prediction Center. The primary changes made to create this dataset were the deletion of some columns, change of some data types, and sorting by date. The period of the data ranges from 1950 to 2021.[^1]

[^1]: source: Diving into US Tornado Data from Kaggle



```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the Data

Download the data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2023-05-16")
```


# Readme

The data contains information about date, year, month and days of Tornados happening in the US.

```{r Readme, eval = interactive()}
tt$tornados%>%head
```


# Wrangle

Selection of the data to produce a map visualization.

```{r Wrangle}

tornados <- tt$tornados %>%
  select(yr,st,slat,slon,elon,elat,len,wid)%>%
  filter(yr=="2022")

tornados%>%head
```


# Visualize

Making a map

```{r Visualize}

tornados%>%
  filter(st=="WI") %>%
  ggplot(aes(x=slon,y=slat,group=st))+
  geom_point(aes(size=wid))
  
```


```{r}
states <- map_data(map = "state")
states %>%head
```

```{r}
states %>%
  ggplot(aes(x=long,y=lat,group=group))+
  geom_polygon(fill="grey",color="white")+
  coord_quickmap()
```


```{r}
wi<- tornados%>%
  filter(st=="WI")


states %>% # count(region)
  ggplot(mapping = aes(x=long,y=lat,group=group))+
  geom_polygon(fill="grey",color="white")+
  geom_polygon(data=states%>%filter(region=="wisconsin"),
               inherit.aes = TRUE,fill="steelblue",alpha=0.5)+
  geom_point(data= wi, 
             mapping=aes(x=slon,y=slat,
                         group=st,
                         size=wid),
             shape=21, stroke=0.5)+
  coord_quickmap()+
  labs(title="Location of 2022 Tornados in WI, US",
       caption = "DataSource: NOAA's National Weather Service Storm Prediction Center",
       size="Width in yards")+
  ggthemes::theme_map(base_size = 16,base_family = "Roboto Condensed")
  
```



# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "images/02_map_Tornados.png",
  device = "png",bg = "white")

```
