---
title: "TidyTuesday - NASA Global Temperature Change - part one"
date: 2023-07-12
output: html_document
---

# Overview


Air temperatures on Earth have been rising since the Industrial Revolution. While natural variability plays some part, the preponderance of evidence indicates that human activities—particularly emissions of heat-trapping greenhouse gases—are mostly responsible for making our planet warmer.

According to an ongoing temperature analysis led by scientists at NASA’s Goddard Institute for Space Studies (GISS), the average global temperature on Earth has increased by at least 1.1° Celsius (1.9° Fahrenheit) since 1880. The majority of the warming has occurred since 1975, at a rate of roughly 0.15 to 0.20°C per decade. [^1]

[^1]: Source: https://earthobservatory.nasa.gov/world-of-change/global-temperatures#:~:text=According%20to%20an%20ongoing%20temperature,1.9%C2%B0%20Fahrenheit)%20since%201880.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE)

library(tidyverse)
library(tidytuesdayR)
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}
tt <- tt_load("2023-07-11")
global_temps <- tt$global_temps
global_temps %>% head
```


# Readme

Data is divided into different variables including Year (1880 - 2023) and months containing the average level of Global Temperatures.

```{r Readme, eval = interactive()}
global_temps %>%names
```


# Wrangle

Here is a exploration of the data to summarise information about Global Temperature change since 1880.

```{r Wrangle}
diff <- global_temps%>%
  select(1:13) %>%
  pivot_longer(cols = -Year)%>%
  #mutate(color=ifelse(value>0,"up","down"))%>%
  group_by(Year)%>%
  reframe(avg_val=mean(value))%>%
  mutate(diff=c(0,diff(avg_val))*100) 

diff%>% head
```

```{r}
diff %>% 
  filter(avg_val > 0)
```



# Visualize

Using your processed dataset, create your unique visualization.

```{r}
tag<-tibble(tag_history= c("The basic GISS temperature analysis scheme was defined in the late 1970s by James Hansen when a method of estimating global temperature change was needed for comparison with one-dimensional global climate models."),
            tag_stats = c("According to an ongoing temperature analysis led by scientists at NASA's Goddard Institute for Space Studies (GISS), the average global temperature on Earth has increased by at least 1.1° Celsius (1.9° Fahrenheit) since 1880."),
            tag_reading =c("How to interpret the background: The dashed-line in the background shows the average Global temperature time series with one month-period lag."))
```

```{r}
global_temps %>%
  select(1:13) %>% # count(Year) 1880 - 2023
  pivot_longer(cols = -Year)%>%
  mutate(color=ifelse(value>0,"up","down"))%>%
  ggplot(aes(x=Year,y=value))+
  geom_col(aes(fill=color))
```
```{r}
ggsave("images/03_first_try.png")
```


```{r Visualize}
library(grid)

global_temps %>%
  select(1:13) %>% # count(Year) 1880 - 2023
  pivot_longer(cols = -Year)%>%
  mutate(color=ifelse(value>0,"up","down"))%>%
  ggplot(aes(x=Year,y=value))+
  geom_line(data=diff,mapping=aes(x=Year,y=diff),
            inherit.aes = F,linetype="dashed",
            color="grey",
            linewidth=0.05)+
  geom_col(aes(fill=color))+
  geom_segment(aes(x=min(Year)-1,xend=min(Year)-1,
                   y=0,yend=-10),color="grey70",
               linewidth=1.5,
               lineend="butt",
               arrow=arrow(length = unit(0.1, "inches")))+
  geom_segment(aes(x=max(Year)+1,xend=max(Year)+1,
                   y=0,yend=10),color="grey70",
                 linewidth=1.5,
                 lineend="butt",
                 arrow=arrow(length = unit(0.1, "inches")))+
  geom_segment(aes(x=1940,xend=1940,
                 y=0,yend=10),color="grey70",
               linewidth=0.5,
               lineend="butt",
               arrow=arrow(length = unit(0.1, "inches")))+
  geom_segment(aes(x=1957,xend=1957,
               y=0,yend=10),color="grey70",
             linewidth=0.5,
             lineend="butt",
             arrow=arrow(length = unit(0.1, "inches")))+
   geom_segment(aes(x=1979,xend=1979,
               y=0,yend=-10),color="grey70",
             linewidth=0.5,
             lineend="butt")+
  ggtext::geom_textbox(data = tag,aes(x=1979,y=-15,label = tag_stats),
                       size = 3, 
                       family="Roboto Condensed",
                       width = unit(20, "line"), 
                       alpha = 0.9,
                       color="grey70",
                       fill="grey4",
                       box.colour = "grey70") +
  ggtext::geom_textbox(data = tag,aes(x=1920,y=-25,label = tag_reading),
                     size = 3, 
                     family="Roboto Condensed",
                     width = unit(20, "line"), 
                     alpha = 0.9,
                     color="grey70",
                     fill="grey4",
                     box.colour = "grey4") +
  geom_hline(yintercept = 0,linewidth=2,color="grey70")+
  scale_x_continuous(n.breaks = 10)+
  scale_y_continuous()+
  annotate(geom = "text",
         family="Roboto Condensed",
         fontface="bold",
         label="Global surface temperatures change\n1880 - 2023",
         size=10,
         color="grey70",
         hjust=0,
         x = 1880 ,y =c(22) )+
   annotate(geom = "text",
         family="Roboto Condensed",
         fontface="bold",
         label="First rise of 14%\nto previous year in 1940 ",
         size=3,
         color="grey70",
         hjust=0,
         x = 1941 ,y =c(13) )+
    annotate(geom = "text",
        family="Roboto Condensed",
        fontface="bold",
        label="Second big rise of 24%\nto previous year in 1957",
        size=3,
        color="grey70",
        hjust=0,
        x = 1959 ,y =c(7) )+
    annotate(geom = "text",
      family="Roboto Condensed",
      fontface="bold",
      label="Steady average rise of 9.3%\nsince 1979",
      size=3,
      color="grey70",
      hjust=0,
      x = 1980 ,y =c(-7) )+
  annotation_custom(grob = grid::circleGrob(x=0,y=0.1,gp=gpar(col="grey70",fill=NA)),
                    xmin = 1940,
                    xmax = 1950,
                    ymin = 0,ymax = 10)+
  ggthemes::scale_fill_fivethirtyeight()+
  labs(title="",
       caption = "\nDataSource: NASA GISS Surface Temperature Analysis (GISTEMP v4)\nDataViz: #TidyTuesday 2023 - week 28 by Federica Gazzelloni\n",
       fill="Temperature",
       y="Monthly Means")+
  #ggthemes::theme_clean()+
  theme_void()+
  theme(text=element_text(color="grey70",family="Roboto Condensed"),
        plot.caption = element_text(hjust = 0.5,lineheight = 1),
        axis.text.x = element_text(color="grey70"),
        plot.background = element_rect(color="grey4",
                                       fill="grey4"),
        legend.position = "bottom",
        legend.title = element_text(color="black"),
        legend.text = element_text(color="black"),
        legend.background = element_rect(color="grey70",fill="grey70"))
```
# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "images/03_tidytuesday_Global_Temps.png",
  device = "png")

```
