---
title: "Time Series Analysis on Global Temperature Anomalies"
date: 2023-07-14
output: html_document
---

# Overview

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
```

# Load the Data

Data is from NASA's Goddard Institute for Space Studies (GISS).

The average global temperature on Earth are estimation of the temperature anomalies valued based on the temperature values in C° taken from 1951 to 1980. It is technically a time series analysis.

```{r Load}
tt <- tt_load("2023-07-11")
```


```{r Data_selection}
global_temps <- tt$global_temps
global_temps%>%head
```

The first step is to set up a dataframe customized as needed for the analysis.

```{r}
global_temps2 <- global_temps %>%
  select(1:13) %>% # count(Year) 1880 - 2023
  pivot_longer(cols = -Year)%>%
  drop_na()%>%
  group_by(Year)%>%
  reframe(avg_val=mean(value))

global_temps2 %>%head
```

Start building a time series dataframe with the help of the {fpp3} package.

```{r}
# install.packages("fpp3")
library(fpp3)
```

```{r eval=FALSE}
?tsibble()
?as_tsibble
```

```{r}
ts_gt <- global_temps2 %>% # names
  as_tsibble(index=Year)

ts_gt %>%head
```

```{r}
autoplot(ts_gt)
```


Apply an ARIMA model.
```{r}
fit <- ts_gt %>%
  model(ARIMA(avg_val~Year))
```


```{r}
report(fit)
```


Let's predict the future:
```{r}
new_data <- new_data(ts_gt,n=5)

fc <- fit |>
  forecast(new_data = new_data)
```

```{r}
fc |>
  autoplot(ts_gt |> tail(100)) +
  labs(title="Global temperature anomalies",
       y = "Temperature variation", x = "Time [Year]")
```

Making a simple linear regression model with our data:
```{r}
global_temps %>%
  select(1:13) %>% # count(Year) 1880 - 2023
  pivot_longer(cols = -Year)%>%
  ggplot(aes(x=Year,y=value,group=name,color=name))+
  geom_line()+
  scale_color_manual(values = RColorBrewer::brewer.pal(12,"Paired"))+
  ggthemes::theme_fivethirtyeight()
```
In order to make credible linear model results, it might be important to split the data into training and testing sets.

```{r}
global_temps_lm_df <- global_temps %>%
  select(1:13) %>% # count(Year) 1880 - 2023
  pivot_longer(cols = -Year)
```

```{r}
set.seed(1234)
train_id <-  sample_frac(tibble(id=row_number(global_temps_lm_df)),0.8)
training <- global_temps_lm_df[pull(train_id),]
testing <-  global_temps_lm_df%>%anti_join(training)
```

```{r}
training%>%head
```


```{r}
fit<- lm(value ~ Year, data=training)
summary(fit, show.intercept= FALSE)
```

```{r}
broom::augment(fit)%>%
  left_join(global_temps_lm_df,by=c("Year","value"))%>%
  ggplot(aes(x=Year,value,group=name))+
  geom_line(color="steelblue",linewidth=0.5)+
  geom_line(aes(y=.fitted),inherit.aes = T)
```

```{r}
new_data <- tibble(Year=c(2024,2025,2026,2027))

predict(fit,new_data)
```
```{r}
prediction<- tibble(Year=c(2024,2025,2026,2027),
                    pred=predict(fit,
                                 newdata = tibble(Year=c(2024,2025,2026,2027))))
prediction
```


```{r}
broom::augment(fit)%>%
  left_join(global_temps_lm_df,by=c("Year","value"))%>%
  ggplot(aes(x=Year,value))+
  geom_line(aes(group=name),color="steelblue",linewidth=0.5)+
  geom_line(aes(y=.fitted),inherit.aes = T)+
  geom_line(data=prediction, mapping=aes(x=Year,y=pred),color="darkred")+
  labs(title="Global Temperature Anomalies Time Series",
  subtitle="Linear model prediction for the next 4 Years",
  caption="DataSource: NASA's Goddard Institute for Space Studies (GISS)")+
  ggthemes::theme_fivethirtyeight()
```

```{r}
ggsave("images/05_ts_linear_model.png")
```

