---
title: "GPT detection models - TidyTuesday 2023 week 29"
date: 2023-07-19
output: html_document
---

# Overview

In this study the authors carried out a series of experiments passing a number of essays to different **GPT detection models**. 

The authors argue that **GPT detectors** disproportionately classify real writing from non-native English writers as AI-generated.


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the Data

Download the data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-07-18")
# tt%>%class
```


# Readme

The detectors dataset is from a R package named detectors, to install the package: 
```{r}
require(pak)
pak::pak("simonpcouch/detectors")
```


```{r Readme, eval = interactive()}
tt
```


# Glimpse Data

The data is composed of 9 variables and 6,185 observations containing the type of detectors used to generate the predictions (**detector**) and the prediction whether is AI producing the text or not (**.pred_AI**), then there is a variable with the specification of the type of model used to generate the essay (**model**).

```{r Glimpse, eval=FALSE,echo=FALSE}

tt %>% 
  map(glimpse)

```

```{r}
detectors <- tt$detectors
```


```{r}
detectors %>% str()
```
```{r}
detectors %>%
  DataExplorer::profile_missing()
```

# Wrangle

Explore the data and process it into a nice format for plotting! 


Let's select the 40% of the original observations which are without missing values.
```{r Wrangle}
detectors_short <- detectors %>%
  select(-prompt) %>%
  drop_na() 

detectors_short %>%head
```

```{r}
detectors_short %>% 
  count(kind,native,.pred_class)%>% 
  group_by(native)%>%
  mutate(prop=round(n/sum(n)*100,2))
```

# Visualize

Using your processed dataset, create your unique visualization.

```{r Visualize}
detectors %>%
  ggplot(aes(.pred_AI))+
  geom_density()
```
```{r Visualize2}
detectors %>%
  ggplot(aes(.pred_AI))+
  geom_histogram()
```


Which type of detectors did wrong?
```{r}
my_df <- detectors_short %>% 
  count(kind,native,.pred_class,detector)%>% 
  group_by(native)%>%
  mutate(prop=round(n/sum(n)*100,2),
         checks=ifelse(kind==.pred_class,"ok","no"))
  
```

```{r}
my_df %>%
  #mutate(native=ifelse(native=="no","Non Native","Native"))+
  ggplot(aes(x=detector,y=prop,group=checks,color=checks))+
  geom_point()+
  geom_line()+
  facet_wrap(vars(native),ncol = 1)+
  labs(title="It very important to be a native English\nto be catched right using AI for essays",caption="DataSouce: #TidyTuesday 2023 week 29 GPT Detectors | DataViz: @fgazzelloni")+
  tvthemes::scale_color_avatar()+
  tvthemes::theme_avatar()+
  theme(plot.caption = element_text(hjust = 0.5))
```


# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "images/08_GPT_detectors.png",
  device = "png")

```
