---
title: "Data Wrangling Best Practices - Historical Markers"
subtitle: "part one"
date: "2023-07-20"
output: html_document
---

# Overview

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2023-07-04")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse)

```

# Wrangle

Explore the data and process it into a nice format for plotting! Access each dataset by name by using a dollarsign after the `tt` object and then the name of the data set.

```{r Wrangle}

historical_markers <- tt$historical_markers
no_markers <- tt$no_markers
```


```{r}
historical_markers%>%count(missing)
```




Make a quick map:
```{r}
states <- map_data("state")
states%>%head
```

```{r}
map <- historical_markers%>%
  filter(!is.na(missing))%>%
  ggplot(aes(x=longitude_minus_w,y=latitude_minus_s))+
  geom_polygon(data=states,mapping=aes(x=long,y=lat,group=group))+
  geom_point(aes(color=missing,fill=missing),
             shape=21,stroke=0.5)+
  ggthemes::theme_map()

map
```


```{r}
historical_markers%>%
  select(title,subtitle,addl_subtitle)%>%head()# %>%View
```

```{r}
historical_markers%>%
  filter(!is.na(subtitle))%>%
  select(title,subtitle,addl_subtitle)%>%
  count(subtitle,sort=T)
```

```{r}
historical_markers%>%
  filter(!is.na(subtitle),!is.na(year_erected))%>%
  select(title,
         subtitle,
         addl_subtitle,
         year_erected,
         state_or_prov)%>%
  count(title,year_erected,state_or_prov,sort=T)%>%
  arrange(year_erected)
```

```{r}
historical_markers%>%
  filter(!is.na(subtitle),!is.na(year_erected))%>%
  select(title,subtitle,addl_subtitle,year_erected,state_or_prov)%>%
  count(title,year_erected,state_or_prov,sort=T)%>%
  ggplot(aes(x=year_erected,y=n))+
  geom_col(aes(color=state_or_prov,fill=state_or_prov),
           alpha=0.5,
           show.legend = F)
```


```{r}
df <- historical_markers%>%
  filter(!is.na(subtitle),!is.na(year_erected))%>%
  select(title,subtitle,addl_subtitle,year_erected,state_or_prov)%>%
  count(title,year_erected,state_or_prov,sort=T)

df %>%
  head
```


```{r}
titles <- df %>%
  select(title)%>%
  mutate(title=tolower(title),
         title = gsub("[[:punct:]]","",title),
         title = gsub("[[:digit:]]","",title)) #%>%
  # sample_frac(0.01) %>% head

titles_ch <- titles$title
```

```{r}
titles_ch[1]
```


```{r}
# words <- str_split(titles," ") 

words <- strsplit(titles_ch, "\\s+")[[1]]
words
```



# Final Visualization

Using your processed dataset, create your unique visualization.

```{r Visualize}

map+
  labs(title="Missing Markers US Map",
       caption = "RStudio Lab DataViz @fgazzelloni | #TidyTuesday 2023 week27")+
  scale_color_viridis_d()+
  theme(plot.title = element_text(size=18),
        plot.caption = element_text(hjust = 0.5))
  
```



# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "images/09_Missing_Markers_map.png",
  device = "png")

```
